---
title: "ABC_PopDynamic_Plateau_Basin"
author: "Kurt M. Wilson & Simon C. Brewer"
date: "2022-10-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the libraries
```{r}
library(viridis)#plotting
library(scales)#plotting
library(ggplot2)#plotting
library(doParallel)#parallel processing
library(doSNOW)#parallel processing
library(utils) #for progress bars
library(rcarbon)#14c date calibration and spd recreation
library(foreach)

source("03_loggrowthmodel.R")
source("04_simulatedmodelspds.R")
source("05_exportbestmodelspds.R")

```


Read in data
```{r}

plateau.14c <- read.csv("./data/Plateau_dates_quality_controlled.csv")

basin.14c <- read.csv("./data/Basin_dates_quality_controlled.csv")

plateau.clim <- read.csv("./data/Plateau.clim.csv")

basin.clim <- read.csv("./data/Basin.clim.csv")

```

Notes about the data
- Radiocarbon dates are those from the region (subset by the the outer boundaries of the watersheds), with any date with an error of > 100 removed. 

First, we will calibrate the radiocarbon dates
```{r calibrate}

#calibrate all of the radiocarbon ages for the plateau
dat_cal.Plat <- calibrate(plateau.14c$Date,#read the age as the 14C value
                        errors = plateau.14c$Error,#the +- standard deviation
                        calCurves = "intcal20"#check wich curve to use (SHCAL20 or MARINE20)
                        )

#repeat for the basin
dat_cal.Basin <- calibrate(basin.14c$Date,#read the age as the 14C value
                        errors = basin.14c$Error,#the +- standard deviation
                        calCurves = "intcal20"#check wich curve to use (SHCAL20 or MARINE20)
                        )
```

To evaluate how much influence there may be in over-representation of individual sites we can check the impact of binning dates at various intervals. Here we are going to check the past 5000 years and an impact of not binning, binning at 0, 500, 1000, 1500, and 2000 year intervals

```{r binsense}



Plat_binsense <- binsense(dat_cal.Plat,plateau.14c$Site_Name, h=seq(0,2000,500), timeRange = c(5000,0), runm = 100, calendar = "BP", binning = "calibrated")#check bins from 0 to 2000 in 500 year intervals using a 100 year running mean.

Basin_binsense <- binsense(dat_cal.Basin,basin.14c$Site_Name, h=seq(0,2000,500), timeRange = c(5000,0), runm = 100, binning = "calibrated", calendar = "BP")#check bins from 0 to 2000 in 500 year intervals using a 100 year running mean.


```

Binsense analyses suggest the greatest changes in SPD occur when little to no clustering occurs (h < 500), particularly for the Great Basin. However clustering with h > 500 produces minimal change. Therefore we elect to bin at 1000 years following Codding et al 2021. Regardless of h value though, the trend in spds remains the same.

Generate the bins we will use in model creation and evaluation
```{r}
Plat.bins <- binPrep(sites = plateau.14c$Site_Name, ages = plateau.14c$Date, h = 1000)
Basin.bins <- binPrep(sites = basin.14c$Site_Name, ages = basin.14c$Date, h = 1000)
```

Using the calibrated dates and the bins, make a summed probability distribution just to see the patterns.
```{r spd non-corrected}

dat_spd.Plat <- spd(dat_cal.Plat, runm = 100, bins = Plat.bins, timeRange = c(5000,0), spdnormalised = FALSE)#make the spd for the plateau dates, we use a running mean of 100 years here for visualization

dat_spd.Basin <- spd(dat_cal.Basin, runm = 100, bins = Basin.bins, timeRange = c(5000,0), spdnormalised = FALSE)#make the spd for the basin dates, we use a running mean of 100 years here for visualization

```

Plot the SPDs
```{r}

cp.col <- rgb(59, 15, 112, max = 255)#purple
gb.col <- rgb(96,206,172, max = 255)#teal


plot(NA,
     xlim=c(0,5000),
     ylim=c(0,2),
     xlab="Year cal. BP",
     ylab="Population Density (SPD)"
)

lines(dat_spd.Plat$grid$PrDens~dat_spd.Plat$grid$calBP, lwd = 1, col = cp.col)

polygon(y=c(rep(0, length(dat_spd.Plat$grid$calBP)), rev(dat_spd.Plat$grid$PrDens)),
        x=c(dat_spd.Plat$grid$calBP, rev(dat_spd.Plat$grid$calBP)),
        col=alpha(cp.col, alpha=0.5),
        border = NA
        )

lines(dat_spd.Basin$grid$PrDens~dat_spd.Basin$grid$calBP, lwd = 1, col = gb.col)


polygon(y=c(rep(0, length(dat_spd.Basin$grid$calBP)), rev(dat_spd.Basin$grid$PrDens)),
        x=c(dat_spd.Basin$grid$calBP, rev(dat_spd.Basin$grid$calBP)),
        col=alpha(gb.col, alpha=0.5),
        border = NA
        )

legend("topright", legend = c("Plateau", "Basin"), fill = c(cp.col, gb.col), border = NA, bty = "n")

```

What we can see is there is apparent population increase in both regions, though that pattern looks quite different between the Plateau and the Basin

However, we know there is going to be error/uncertainty around these SPD recreations of relative past population density. This is important because the SPD is our observed statistic for model matching. Therefore we don't want to match our simulated spd just to the spd plotted above. To better capture the uncertainty in the spd recreations of relative past population density, and to better enable us to evaluate model parameter sets' performace, we will implement an adapted version Dinapoli et al's 2021 generation of a thinned spd. Thinning causes us to use a single date per bin as the 'true' date. Dinapoli et al generated a new thinned spd for each run of their model. We will instead generate 1000 unique spds (via iterated spd creation with thinning). This enables several key things: 1) this will help capture the uncertainty in the generated spd, 2) by calculating a simulated model's fit to each of the 1000 spds we can weight selection of high performing models towards those that best match to the full spread of the observed spd with its uncertainty rather than single parameter sets that may match a single spd well. For example, say one model fits really well to 75% of our observed spds - in our setup by using 1000 spds we can ensure more value is assigned to that model than one that only matches 25% of spds well.


First, make normalised and non-normalised calibrated date objects. We use both normalised and not normalised due to uncertainty as to which is the most appropriate choice (see Dinapoli et al 2021).
```{r}
#Calibrate the dates using normalisation
dat_cal.Plat.norm <- calibrate(plateau.14c$Date, #read the 14c age
                          errors = plateau.14c$Error, #get the date's error
                          calCurves = "intcal20",  #use Intcal 20 to calibrate the 14c dates
                          normalised = TRUE)

#and without normalising
dat_cal.Plat.nnorm <- calibrate(plateau.14c$Date,
                                errors = plateau.14c$Error,
                                calCurves = "intcal20",
                                normalised = FALSE)
```


To begin matching simulations to our observed SPD data, we first begin by defining target spds.
```{r}

#Using a thinning process so we use only 1 date per generated bin (from above)

thin.df <- data.frame(RN = 1:length(Plat.bins), bins = Plat.bins)#First, make a dataframe with a unique numeric identifier for each date
allbins <- unique(as.character(Plat.bins)) #grab the unique bins (unike sites)
sitelist <- vector(mode = "list", length = length(allbins))#make a vector of lists that is equal to the number of unique bins

#Set how many spds we will build for comparing
n.spds <- 1000

#setup for parallel processing & tracking progress
ncores <- 20
cl <- makeCluster(ncores)
registerDoSNOW(cl)
pb <- txtProgressBar(max = n.spds, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

results <- data.frame()
#class(results) <- 'CalGrid'

observed.spds <- foreach(i = 1:n.spds, .packages = "rcarbon", .options.snow = opts) %dopar%
  {
  set.seed(i)
  #randomly thin the 14c dates by bins to get a subsampled set of dates with 1 date per bin.
  for (a in 1:length(allbins)) { 
  thin.df1 <- thin.df[thin.df$bins == allbins[a],]#grab each of the unique bins and their numeric identifier
  if (nrow(thin.df1)<=1){
    sitelist[[a]] <- thin.df1$RN #if there is only 1 date from a bin, use that bin's numeric identifier 
  } else {
    sitelist[[a]] <- sample(thin.df1$RN, 1)#if there is more than 1 date from a bin, sample the bin and keep only 1 of the unique identifiers
  }
}

index = sort(unlist(sitelist))#get the sorted numeric identifiers for each bin selected above as a vector

#Create a target SPD (normalised) that uses only the number of dates equal to (and positioned equal to) those selected in the thinning process (1 date per bin)
target.spd.norm <- spd(dat_cal.Plat.norm[index], datenormalised = FALSE, spdnormalised = TRUE, timeRange = c(5000,0), verbose = FALSE)
#And repeat for a non-normalised spd (non-normalised b/c it uses the non-normalised calibrated date object)
target.spd.nnorm <- spd(dat_cal.Plat.nnorm[index], datenormalised = FALSE, spdnormalised = TRUE, timeRange = c(5000,0), verbose = FALSE)

return(data.frame(calBP = target.spd.norm$grid$calBP, norm.PrDens = target.spd.norm$grid$PrDens, nnorm.PrDens = target.spd.nnorm$grid$PrDens))
    
  }
    
stopCluster(cl)

observed.spds <- do.call('cbind.data.frame', observed.spds) #Bind the results together
observed.spds <- observed.spds[, -seq(from = 4,to = length(observed.spds[1,]), by = 3)]#drop the repeated yBP values



```
Save these to a csv so the code does not have to be rerun
```{r}

#write.csv(observed.spds, "Observed_spds_1000reps.csv")

observed.spds <- read.csv("./data/Observed_spds_1000reps.csv")

observed.spds.norm <- observed.spds[,c(1,seq(2,2002,2))]#grab the normalized spds and put in own object
observed.spds.nnorm <- observed.spds[,c(1,seq(3,2003,2))]#make a df for the not normalized spds
```

Plot the observed spds
```{r}

par(mfrow = c(1,2))
plot(observed.spds.norm$norm.PrDens~observed.spds.norm$calBP, type = "l", col = alpha("grey50", alpha = 0.25), ylab = "Summed Probability (Normalized)", xlab = "CalBP", ylim = c(0, 0.0008))
for (i in 2:1002) {
  lines(observed.spds.norm[,i]~observed.spds.norm$calBP, col = alpha("grey50", alpha = 0.25))
}

plot(observed.spds.nnorm$nnorm.PrDens~observed.spds.nnorm$calBP, type = "l", col = alpha("grey50", alpha = 0.25), ylab = "Summed Probability (Normalized)", xlab = "CalBP", ylim = c(0,0.0008))
for (i in 2:1002) {
  lines(observed.spds.nnorm[,i]~observed.spds.nnorm$calBP, col = alpha("grey50", alpha = 0.25))
}

```

Here we can see the outcome of 1000 iterations of each the normalised and not normalised observed spd. This gives us 1000 possible recreations of the 'true' spd. As we can see, the shape and direction does not change between the samples, but there is variation in the band. Thus we will match to the range of this variation.

Next we can visualize the environmental trends over the same period.
Plot the averaged precip per month for each of the regions
```{r}
plot(plateau.clim$MoAvgPrecip~plateau.clim$calBP, type = "l", lwd = 2, col = cp.col, ylim = c(22, 34), ylab = "Avg. Mo. Precip.", xlab = "CalBP")
lines(basin.clim$MoAvgPrecip~basin.clim$calBP, lwd = 2, col = gb.col)
legend("bottomright", legend = c("Plateau", "Basin"), col = c(cp.col, gb.col), lty = 1, cex = 0.75, bty = "n")

```

Plot average Tmin and Tmax per region
```{r}
par(mfrow = c(1,2))

plot(plateau.clim$MoAvgTmax~plateau.clim$calBP, type = "l", lwd = 2, col = cp.col, ylim = c(13, 19), ylab = "Avg. Mo. Tmax", xlab = "CalBP")
lines(basin.clim$MoAvgTmax~basin.clim$calBP, lwd = 2, col = gb.col)
legend("topright", legend = c("Plateau", "Basin"), col = c(cp.col, gb.col), lty = 1, cex = 0.75, bty = "n")

plot(plateau.clim$MoAvgTmin~plateau.clim$calBP, type = "l", lwd = 2, col = cp.col, ylim = c(-4, 2), ylab = "Avg. Mo. Tmin", xlab = "CalBP")
lines(basin.clim$MoAvgTmin~basin.clim$calBP, lwd = 2, col = gb.col)
legend("topright", legend = c("Plateau", "Basin"), col = c(cp.col, gb.col), lty = 1, cex = 0.75, bty = "n")


```

##NEED TO EDIT THIS FOR UPDATED NPP ONCE WE CALCULATE THAT WITH THE NEW TEMP AND PRECIP VALUES
Read in the climate data for model development
```{r}
clim.df <- read.csv("./data/clim.df.csv")
```


Establish parameter combinations
```{r}

nsim <- 100 #set the number of parameter combinations to use

npp_lo <- vector()#a vector to hold the npp min threshold for any population to be present
npp_hi <- vector()#a vector to hold the npp max threshold over which no increase in k occurs (unless by intensification)
for (i in 1:nsim) {
  npp_limits <- sort(runif(2, min = 0.1, max = 0.8))#grab 2 npp values to be low and high and sort from low to high
  npp_lo <- append(npp_lo, npp_limits[1])  
  npp_hi <- append(npp_hi, npp_limits[2])
}
tstart <- round(runif(nsim, min = 3000, max = 5000))#select the year in which intensification will begin
cK0 <- round(runif(nsim, min = 2, max = 10))#set a starting population
r <- exp(runif(nsim, min = log(0.01), max = log(0.5)))#set a population growth rate
Kmax <- exp(runif(nsim, min = log(100), max = log(10000)))#set a maximum carrying capacity

#make a dataframe to hold nsim parameter combinations
params <- data.frame(tstart = tstart, r = r, cK0 = cK0, Kmax = Kmax, npp_lo = npp_lo, npp_hi = npp_hi)
params$combo <- 1:length(params$tstart)#set a unique parameter combination number so we can link outcomes to initial param values


```

Visualize the npp to k relationship
```{r}

plot.df <- data.frame(npp = c(min(clim.df$npp), params$npp_lo[1], params$npp_hi[1], max(clim.df$npp)),
                      K = c(0, 0, params$Kmax[1], params$Kmax[1]))
ggplot(plot.df, aes(x = npp, y = K)) +
  geom_line()

```

Run the model and generate the distance measure outputs
```{r}
#setup for parallel processing & tracking progress
ncores <- 10
cl <- makeCluster(ncores)
registerDoSNOW(cl)
pb <- txtProgressBar(max = nsim, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

g.model.res <- foreach(i = 1:nsim, .packages = "rcarbon", .options.snow = opts) %dopar%
  {

    #establish npp boundary conditions
    npp.conds <- data.frame(npp = c(params$npp_lo[i], params$npp_hi[i], max(clim.df$npp)),
                      K = c(0, params$Kmax[i], params$Kmax[i])) #dataframe where k at npp_lo = 0 (min threshold for population survival - anything at or below this will have a k of 0), at npp_hi = max k, and at highest observed npp = maxk - this is all used below for the npp to k relationship for the model run
    
    ## Derive time-dependent K
    ## this will give us k at each time step (or at least k without any impact of SEI - so k based on npp)
    ## First filter climate to tstart
    clim.df <- clim.df[clim.df$age < tstart[i],] #
    time <- max(clim.df$age) - clim.df$age## Need to invert time somehow
    K_t <- approx(npp.conds$npp, npp.conds$K, clim.df$npp)$y #get a list of interpolated points connecting npp and k at each observation of npp     from our climate reconstruction
    #plot(clim.df$age, K_t)
    
    set.seed(i)
    sim.growthModel <- data.frame(log_growth_t(N0 = cK0[i], r = r[i], K = K_t, time = time))
    sim.growthModel[,c(1,4)]
    class(sim.growthModel) <- 'CalGrid'
    param.combo <- params$combo[i] #hold onto the unique parameter combo
    results <- createmodelspd(model = sim.growthModel)#create simulated spd from simulate growth model, need to use just calBP and PrDens
    return(results)
    
    #this line of code would be used to do the direct Dinapoli et al approach
    #sim.growthModel <- log_growth_t(N0 = cK0[i], r = r[i], K = K_t, time = time)

  }


stopCluster(cl)

eval.results<- do.call('rbind.data.frame', g.model.res)

```
Select the top performing models
```{r}
no.models.keep <- 5

#get the top 5 performing models from each measure
top.params.eucun <- eval.results[order(eval.results$euc.uncal.norm),][1:no.models.keep,]
top.params.eucunn <- eval.results[order(eval.results$euc.uncal.nnorm),][1:no.models.keep,]
top.params.ecun <- eval.results[order(eval.results$euc.cal.norm),][1:no.models.keep,]
top.params.ecunn <- eval.results[order(eval.results$euc.cal.nnorm),][1:no.models.keep,]
top.params.nrmseun <- eval.results[order(eval.results$nrmse.uncal.norm),][1:no.models.keep,]
top.params.nrmseunn <- eval.results[order(eval.results$nrmse.uncal.nnorm),][1:no.models.keep,]
top.params.nrmsecn <- eval.results[order(eval.results$nrmse.cal.norm),][1:no.models.keep,]
top.params.nrmsecnn <- eval.results[order(eval.results$nrmse.cal.nnorm),][1:no.models.keep,]

```

We can then rerun the model for just the best performing parameter combinations
```{r}

#first, select the parameter combinations of interest
best.params <- params[params$combo %in% top.params.ecun$param.combo,]
best.params <- rbind(best.params, params[params$combo %in% top.params.eucunn$param.combo,], params[params$combo %in% top.params.ecun$param.combo,],params[params$combo %in% top.params.ecunn$param.combo,],params[params$combo %in% top.params.nrmseun$param.combo,],params[params$combo %in% top.params.nrmseunn$param.combo,],params[params$combo %in% top.params.nrmsecn$param.combo,],params[params$combo %in% top.params.nrmsecnn$param.combo,])


#setup for parallel processing & tracking progress
ncores <- 10
cl <- makeCluster(ncores)
registerDoSNOW(cl)
pb <- txtProgressBar(max = nsim, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

best.models.res <- foreach(i = 1:length(best.params[,1]), .options.snow = opts) %dopar%
  {

    #establish npp boundary conditions
    npp.conds <- data.frame(npp = c(best.params$npp_lo[i], best.params$npp_hi[i], max(clim.df$npp)),
                      K = c(0, best.params$Kmax[i], best.params$Kmax[i])) #dataframe where k at npp_lo = 0 (min threshold for population survival - anything at or below this will have a k of 0), at npp_hi = max k, and at highest observed npp = maxk - this is all used below for the npp to k relationship for the model run
    
    ## Derive time-dependent K
    ## this will give us k at each time step (or at least k without any impact of SEI - so k based on npp)
    ## First filter climate to tstart
    clim.df <- clim.df[clim.df$age < tstart[i],] #
    time <- max(clim.df$age) - clim.df$age## Need to invert time somehow
    K_t <- approx(npp.conds$npp, npp.conds$K, clim.df$npp)$y #get a list of interpolated points connecting npp and k at each observation of npp     from our climate reconstruction
    #plot(clim.df$age, K_t)
    
    set.seed(i)
    sim.growthModel <- data.frame(log_growth_t(N0 = cK0[i], r = r[i], K = K_t, time = time))
    
    #this line of code would be used to do the direct Dinapoli et al approach
    #sim.growthModel <- log_growth_t(N0 = cK0[i], r = r[i], K = K_t, time = time)

    sim.growthModel$combo <- best.params$combo[i]
    #results <- rbind(results, sim.growthModel)
    return(sim.growthModel)
  }


stopCluster(cl)

best.results<- do.call('rbind.data.frame', best.models.res)

```

We can then plot a population size estimate and a PrDens for each year in the simulation from the best performing models.
```{r}
#make a blank plot with xlim as c(0, max(results$time) and ylim as c(0, max(results$N)))

#then do a for loop to plot the lines for N~time for each model

plot(NULL,xlim = c(0, max(best.results$CalBP)), ylim = c(0, max(best.results$PrDens, na.rm = TRUE)), ylab = "Pop as PrDens", xlab = "CalBP")
for (i in 1:nsim){
  lines(subset(best.results$PrDens, best.results$combo == i) ~ subset(best.results$CalBP, best.results$combo == i), lwd = 1, col = alpha("grey50", alpha = 0.5))
}

plot(NULL,xlim = c(0, max(best.results$CalBP)), ylim = c(0, max(best.results$N, na.rm = TRUE)), ylab = "Total Population Size", xlab = "CalBP")
for (i in 1:nsim){
  lines(subset(best.results$N, best.results$combo == i) ~ subset(best.results$CalBP, best.results$combo == i), lwd = 1, col = alpha("grey50", alpha = 0.5))
}

```

We can also plot the top performing models for each distance metric
```{r}
#euc uncal norm
plot(NULL,xlim = c(0, max(best.results$CalBP)), ylim = c(0, max(best.results$PrDens, na.rm = TRUE)), ylab = "Population Size", xlab = "CalBP", main = "EUC UNCAL NORM")
for (i in 1:no.models.keep){
  lines(subset(best.results$PrDens, best.results$combo == i) ~ subset(best.results$CalBP, best.results$combo == i), lwd = 1, col = alpha("grey50", alpha = 0.5))
}

#euc uncal nnorm
plot(NULL,xlim = c(0, max(best.results$CalBP)), ylim = c(0, max(best.results$PrDens, na.rm = TRUE)), ylab = "Population Size", xlab = "CalBP", main = "EUC UNCAL NNORM")
for (i in (no.models.keep+1):(no.models.keep*2)){
  lines(subset(best.results$PrDens, best.results$combo == i) ~ subset(best.results$CalBP, best.results$combo == i), lwd = 1, col = alpha("grey50", alpha = 0.5))
}

#euc cal norm
plot(NULL,xlim = c(0, max(best.results$CalBP)), ylim = c(0, max(best.results$PrDens, na.rm = TRUE)), ylab = "Population Size", xlab = "CalBP", main = "EUC CAL NORM")
for (i in (no.models.keep*2+1):(no.models.keep*3)){
  lines(subset(best.results$PrDens, best.results$combo == i) ~ subset(best.results$CalBP, best.results$combo == i), lwd = 1, col = alpha("grey50", alpha = 0.5))
}

#euc cal nnorm
plot(NULL,xlim = c(0, max(best.results$CalBP)), ylim = c(0, max(best.results$PrDens, na.rm = TRUE)), ylab = "Population Size", xlab = "CalBP", main = "EUC CAL NNORM")
for (i in (no.models.keep*3+1):(no.models.keep*4)){
  lines(subset(best.results$PrDens, best.results$combo == i) ~ subset(best.results$CalBP, best.results$combo == i), lwd = 1, col = alpha("grey50", alpha = 0.5))
}

#NRSME uncal norm
plot(NULL,xlim = c(0, max(best.results$CalBP)), ylim = c(0, max(best.results$PrDens, na.rm = TRUE)), ylab = "Population Size", xlab = "CalBP", main = "NRSME UNCAL NORM")
for (i in (no.models.keep*4+1):(no.models.keep*5)){
  lines(subset(best.results$PrDens, best.results$combo == i) ~ subset(best.results$CalBP, best.results$combo == i), lwd = 1, col = alpha("grey50", alpha = 0.5))
}

#NRSME uncal nnorm
plot(NULL,xlim = c(0, max(best.results$CalBP)), ylim = c(0, max(best.results$PrDens, na.rm = TRUE)), ylab = "Population Size", xlab = "CalBP", main = "NRSME UNCAL NNORM")
for (i in (no.models.keep*5+1):(no.models.keep*6)){
  lines(subset(best.results$PrDens, best.results$combo == i) ~ subset(best.results$CalBP, best.results$combo == i), lwd = 1, col = alpha("grey50", alpha = 0.5))
}

#NRSME cal norm
plot(NULL,xlim = c(0, max(best.results$CalBP)), ylim = c(0, max(best.results$PrDens, na.rm = TRUE)), ylab = "Population Size", xlab = "CalBP", main = "NRSME CAL NORM")
for (i in (no.models.keep*6+1):(no.models.keep*7)){
  lines(subset(best.results$PrDens, best.results$combo == i) ~ subset(best.results$CalBP, best.results$combo == i), lwd = 1, col = alpha("grey50", alpha = 0.5))
}

#NRSME cal nnorm
plot(NULL,xlim = c(0, max(best.results$CalBP)), ylim = c(0, max(best.results$PrDens, na.rm = TRUE)), ylab = "Population Size", xlab = "CalBP", main = "NRSME CAL NNORM")
for (i in (no.models.keep*7+1):(no.models.keep*8)){
  lines(subset(best.results$PrDens, best.results$combo == i) ~ subset(best.results$CalBP, best.results$combo == i), lwd = 1, col = alpha("grey50", alpha = 0.5))
}
    

```


Next, with the simulated model, we need to generate a summed probability distribution. The code below (along with the function in the simulated modelspds source file) allow us to do this. The spd recreation code is mostly replicated from Dinapoli et al 2021.
```{r}
#setup for parallel processing & tracking progress
ncores <- 10
cl <- makeCluster(ncores)
registerDoSNOW(cl)
pb <- txtProgressBar(max = length(best.params[,1]), style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

#Add code to ignore if the model PrDens goes to NA at any point (which happens if we drop to a population of 0)
best.sims.spds <- foreach (i=1:length(best.params[,1]), .packages = c("rcarbon"), .options.snow = opts) %dopar%
  {
    #subset to a single model simulation run
    model <- best.results[which(best.results$combo == best.params$combo[i]),]
    param.combo <- unique(model$combo) #hold onto the unique parameter combo
    model <- model[,c(1,4)]#grab just calbp and prdens from the model
    #convert the model dataframe into a CalGrid object
    class(model) <- 'CalGrid'

#setup a unique identifier
idCounter <- 0

set.seed(i)
    model.spds <- createbestmodelspd(model=model)
    return(model.spds)
    
  }

stopCluster(cl)

```


We can then plot the log growth model spds to see how they look.
```{r}
plot(sim.spds[[1]]$ucs.norm.prdens ~ sim.spds[[1]]$calBP, type = "l", ylab = "PrDens", xlab = "Uncalibrate, Normalised")
plot(sim.spds[[1]]$ucs.nnorm.prdens ~ sim.spds[[1]]$calBP, type = "l", ylab = "PrDens", xlab = "Uncalibrate, Not Normalised")
plot(sim.spds[[1]]$cs.norm.prdens ~ sim.spds[[1]]$calBP, type = "l", ylab = "PrDens", xlab = "Calibrate, Normalised")
plot(sim.spds[[1]]$cs.nnorm.prdens ~ sim.spds[[1]]$calBP, type = "l", ylab = "PrDens", xlab = "Calibrate, Not Normalised")

```
